---
title: "Milestone 1"
author: "Carleena Ortega and Saelin Bjornson"
date: "01/03/2020"
always_allow_html: true
output: 
  html_document:
    toc: true
    keep_md: yes
---

```{r, include=FALSE}
library(ggplot2)
library(here)
library(tidyverse)
library(reshape2)
library(DT)
library(glue)
```

# Task 1: Choosing a dataset
We choose the [Adult Income](https://archive.ics.uci.edu/ml/datasets/adult) data set to analyze for the group project. 

# Task 2. Project Proposal and EDA
## 2.1 Introduce and describe your dataset

Who: The data set was extracted by Barry Becker from the 1994 Census database and is donated by Silicon Graphics \
What: This is a multivariate dataset with categorical and integer variables. It contains the predicted income of individuals from the census with attributes including age, marital status, work class, education, sex, and race. \
When: The data is from a 1994 census. \
Why: The data set is found in the University of California Irvine Machine Learning Repository, and was used for ML prediction of whether a person makes over or under 50K a year based on their attributes. \
How: The census data was collected by survey. 



|Variable|Type|Description|
|--------|-------|------|
|age|int|Age of individual|
|workclass|chr|e.g. private, self-emplowed, federal government, never worked, etc.|
|fnlwgt|int|Final weights: weighted sums of the socio-economic characteristics of the individual. People with similar demographics have similar weights. |
|education|chr|Highest education recieved|
|educationnum|factor|Numerical code for highest education recieved|
|marital_status|int|e.g. married, never married, divorced, etc.|
|occupation|chr|Occupation of individual|
|relationship|chr|Relation of individual in family. e.g. wife, child, husband, unmarried|
|race|chr|Asian-Pacific Islander, Native American, White, Black, other|
|sex|chr|Male or Female|
|capital_gain|int|Profit from capital assets such as investments, real estate, etc.|
|capital_loss|int|Loss from capital assets|
|hours_per_week||Hours the individual works per week|
|country|chr|Country of origin|
|income|chr|Whether individual is predicted to make over or under 50K|





## Task 2.2: Load your dataset (from a file or URL).
```{r, warning=FALSE, message=FALSE}
adult_income <- read_csv("adult.data.csv", col_names=FALSE)
```

```{r, warning=FALSE, message=FALSE}
#To add appropriate column names and change "?" to "NA", and change eduction number to factor

data <- adult_income %>% rename("age"=X1,"workclass"=X2,"fnlwgt"=X3,"education"=X4,
                        "education_num"=X5,"marital_status"=X6,"occupation"=X7,"relationship"=X8,
                        "race"=X9,"sex"=X10,"capital_gain"=X11,"capital_loss"=X12,"hours_per_week"=X13,
                        "country"=X14,"income"=X15) %>%    #give appropriate column names
  mutate(income=recode(income, ">50K"="over_50K", "<=50K"="under_50K")) %>%
  mutate_at(vars("workclass","education","marital_status","occupation","relationship","race","sex","country"), na_if, "?")  
data$education_num <-as.factor(data$education_num)
```

```{r}
hours_age<-lm(hours_per_week~age,data) 
  saveRDS(hours_age, file = glue::glue("hours_age.rds"))

  hours_age
  
```


```{r data tibble,  warning=FALSE, message=FALSE}
head(as.tibble(data))
```


## Task 2.3: Explore your dataset
Perform some exploratory data analysis (EDA) to understand your dataset better. 


**How many variables are present?**
```{r,  warning=FALSE, message=FALSE}
ncol(data)
nrow(data)
```
There are 15 variables, and 32461 observations (people) 



It shows that there are 15 columns, but only 14 variables. And these variables are as follows (education and education_num are similar):
```{r}
colnames(data)
```

There are 15 variables, and 32461 observations (people) 

**What is the range of values for each numerical variable?** 
_Note: Education_Number is a factor, as it is a numerical code for education. But here will change to integer first and treat as integer_
```{r continuous summary,  warnings=FALSE}
sum_df <-data.frame(Age=integer(),Education=integer(),Hours=integer(),Capital_Gain=integer(),Capital_Loss=integer())
means <-summarise(data, mean_age=mean(age), mean_education_level = mean(as.integer(education_num)), mean_hoursPerWeek = mean(hours_per_week)) %>% round(2)
max <- summarise(data, max_age=max(age), max_education_level = max(as.integer(education_num)), max_hoursPerWeek = max(hours_per_week))  %>% round(2)
min <- summarise(data, min_age=min(age), min_education_level = min(as.integer(education_num)), min_hoursPerWeek = min(hours_per_week))  %>% round(2)
std_dev <- summarise(data, sd_age=sd(age), sd_education_level = sd(as.integer(education_num)), sd_hoursPerWeek = sd(hours_per_week))  %>% round(2)
NAs <- data %>% select(age, education_num,hours_per_week,capital_gain,capital_loss) %>% sapply(function(x) sum(length(which(is.na(x)))))
sum_df[1,] <- means[1,]
sum_df[2,] <- max[1,]
sum_df[3,] <- min[1,]
sum_df[4,] <- std_dev[1,]
sum_df[5,] <-NAs
rownames(sum_df) <- c("Mean","Max", "Min", "Standard Deviation", "Number NAs")

sum_df
```



**How many unqiue values and NAs in each categorical variable?**
```{r categorical summary}
categorical_df <-data.frame(Work_Class=integer(),Marital_Status=integer(),Occupation=integer(),Relationship=integer(),Race=integer(),Country=integer())

Unique <-summarise(data, unique_work=length(unique(na.omit(age))), unique_MS=length(unique(na.omit(marital_status))), unique_occupation=length(unique(na.omit(occupation))), unique_relationships=length(unique(na.omit(relationship))), unique_race=length(unique(na.omit(race))), unique_country=length(unique(na.omit(country)))) %>% as.data.frame()

NAs <- data %>% select(workclass, marital_status, occupation, relationship, race, country) %>% sapply(function(x) sum(length(which(is.na(x)))))

categorical_df[1,] <- Unique[1,]
categorical_df[2,] <- NAs
rownames(categorical_df) <- c("Unqiue Values", "Number NAs")

categorical_df
```



Plotting Data

**Distribution of age for men and women:**
```{r Plot1, warning=FALSE, message=FALSE}
df <-select(data, sex, age) %>% mutate(sex = factor(sex, levels=c("Male", "Female"))) %>% group_by(sex)  %>% summarize(mean=mean(age))


data %>% mutate(sex = factor(sex, levels=c("Male", "Female"))) %>%
  ggplot(aes(x=age,fill=sex)) + 
  geom_histogram(alpha=0.8,position="identity") +
  scale_fill_manual(values=c("skyblue2","deeppink4")) + 
  labs(title="Distribution of Age by Sex",x="Age",y="Count") +
  theme_bw() +
  theme(legend.title=element_blank())
```

**Proportion of people making >50K a year for men and women, by race:**
```{r Plot2, warning=FALSE, message=FALSE}
data_2 <- data
data_2$education_num <- as.integer(data_2$education_num)
df <- data_2 %>% select(sex, education_num,income,race)
df$educ[df$education_num < 10] <- "PS"  #for Post-secondary
df$educ[df$education_num >= 10] <- "HS"  #for High School, all values lower than or equal to 10

df %>% filter(income =="over_50K") %>% select(race,educ,sex) %>% group_by(race,educ,sex) %>% tally() %>%
  ggplot(aes(sex, n, fill=educ)) +
  geom_bar(position="fill", stat="identity") +
  theme_bw() +
  scale_fill_manual(values=c("skyblue2","deeppink4"),labels=c("High School","Post-Secondary")) + 
  facet_wrap(~race) +
  labs(title="Education level of people making over 50K",fill="Education",y="Percent",x="Sex")
```



**Hours worked per week by age groups:**
```{r Plot3, warning=FALSE, message=FALSE}
data_3 <- select(data, age, hours_per_week) %>% mutate(Age_level= case_when(age <20 ~ "Under 20",
                                                                       age <30 ~"30s",
                                                                       age <40 ~"40s",
                                                                       age <50 ~"50s",
                                                                       age <60 ~"60s",
                                                                       age <70 ~"70s",
                                                                       age <80 ~"80s",
                                                                       TRUE ~ "90+")) 
data_3 %>% mutate(Age_level =fct_relevel(Age_level,"Under 20")) %>%
  ggplot(aes(Age_level, hours_per_week,fill=Age_level)) +
  geom_boxplot(outlier.size=0.2) +
  labs(title="Hours worked per week by age", x="Age", y="Hours per Week") +
  theme_bw() +
  theme(legend.position="none")

```





**The number of hours worked based on marital status grouped by race:**
```{r Plot4, warning=FALSE, message=FALSE}
data %>% 
  select(hours_per_week,race,marital_status)%>%
  mutate(marital_status = case_when( 
    marital_status == c("Married-AF-spouse","Married-civ-spouse","Married-spouse-absent") ~ "Married",
    TRUE ~ "Single")) %>%
  ggplot()+
  geom_boxplot(aes(marital_status,hours_per_week), outlier.size=0.2) +
  labs(x="Marital Status",y="Hours Worked per Week",
    title="The Relationship between Marital Status and Work Hours")+
  facet_wrap(~race)+ 
  theme_bw()
```


## Task 2.4: Research question & plan of action

In this analysis, we seek to determine the difference in socioeconomic factors such as age, education, sex, and marital status between individuals earning less than and those earning more than $50,000 a year.

**Research Questions and Plan** 



* Logistic regression using all the above categorical and continuous explanatory variables. 

* Plots showing the relationship between income and each variable separately. For example, we will perform a logistic regression to show the the difference between individuals earning more than 50,000 a year and those who don't using the educational level as the independent variable.


2. Is hours worked per week correlated with age, relationship, education level, or sex?

* Linear regression using all the above categorical and continuous explanatory variables.

* Plots showing the relationship between hours worked and each variable separately. For example, we will use the linear regression model to explore how hours at work is related to variables such as age, relationship, education level, and sex.




